\textit{Supporting Information for}: Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-02-01 paper/supporting_information.tex

UMAP achieves a balance between representing local and global structure via a subset of its hyper-parameters: \texttt{n\_neighbors} , \texttt{spread}, and \texttt{min\_dist}.  The \texttt{n\_neighbors} hyper-parameter ($K$) denotes the number of nearest neighbors to consider in constructing the high-dimensional fuzzy simplicial set for each datapoint.  \texttt{Spread} ($\gamma$) and \texttt{min\_dist} ($\delta$) function together to create the differentiable decay curve used to approximate the injective function used to map between high- and low-dimensional fuzzy simplicial sets.  In essence, \texttt{min\_dist} determines the degree to which nearby points are clustered or expanded, relative to to the overall \texttt{spread}.

Two other parameters ultimately affect this balance between preserving local and global structure: the seed ($\tau$) for the (pseudo-)random number generator, and the order of the observations (i.e., trajectories) to be embedded ($\phi$).  As described in \textit{Methods}, we ensured the video and recall events were projected onto the \textit{same} low-dimensional manifold by fitting the embedding model to a stacked matrix of all video, average recall, and individuals' recall events.  After initializing the low-dimensional simplicial set (by default, using a spectral embedding of the high-dimensional simplicial set's fuzzy 1-skeleton), UMAP optimizes the embedding using stochastic gradient descent with cross-entropy as a cost function.  During optimization, indices of the data are sampled at random, and thus local minimum reached by the optimization is dependent on both the state of the RNG and the sequence in which observations are concatenated.

We performed a grid search over pre-specified values of these hyper-parameters, and optimized the manifold space for visualization based on two criteria: First, that the 2D embedding of the video trajectory should reflect its original 100-dimensional structure as faithfully as possible. Second, that the path traversed by the embedded video trajectory should intersect itself a minimal number of times.  The first criteria helps bolster the validity of visual intuitions about relationships between sections of video content, based on their locations in the embedding space.  The second criteria was motivated by the observed low off-diagonal values in the video trajectory's temporal correlation matrix (suggesting that the same topic-space coordinates should not be revisited; see Figure~2A in the main text) and the inability to accurately describe the consistency of participants' transitions through topic space at locations where those transitions frequently occurred in multiple, orthogonal directions (see \textit{Methods} for analysis details).  We formalized this optimization as the combination of hyper-parameter values that satisfied
\[
\argmax_{\left\{K, \gamma, \delta, \tau, \varphi~\in~E~\mathrm{|}~\Phi\left(K, \gamma, \delta, \tau, \varphi \right) \right\}} \left[\mathrm{corr}\left(\Upsilon\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right),~\Upsilon\left(\chi\right)\right)\right],~\mathrm{where}
\]
\[
\Phi\left(K, \gamma, \delta, \tau, \varphi \right) = \argmin_{K, \gamma, \delta, \tau, \varphi} \left[\Gamma\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right)\right],
\]
$\xi$ is the video's trajectory through the manifold space; $\chi$ is the original, 100-dimensional video trajectory; $\Upsilon$ is a condensed matrix of pairwise distances between event vectors (computed using correlation distance in the original topic space and Euclidean distance in the manifold space); $\mathrm{corr}\left(\Upsilon\left(\xi\right), \Upsilon\left(\chi\right)\right)$ is the correlation between the sets of pairwise distances, and $\Gamma$ is the number of instances in which lines drawn between consecutive video event embeddings intersected each other.   The sets of hyper-parameter values we searched over comprised: $K \in \{10x \in \mathbb{Z}~|~10 < x <  22\} \cup \{161\}$ (a range roughly centered on half the total number of events, 161, in order to balance representations of local and global structure); $\gamma \in \{1, 3, 5, 7, 9\}$; $\delta \in \{0.1, 0.3, 0.5, 0.7\}$; $\tau \in \{x \in \mathbb{Z}~|~0 < x < 101\}$; and $\phi \in {S\choose3},~\mathrm{where}~S=\{\mathrm{video~events,~average~recall~events, individual~recall~events}\}$.  The optimal parameters (that yielded $\Phi=0$) were $K=170$, $\gamma=7$, $\delta=0.7$, $\tau=41$, and $\varphi$ as a the sequences of average recall events, video events, and individuals' recall events, vertically concatenated.

>>>>

UMAP achieves a balance between representing local and global structure via a subset of its hyper-parameters: \texttt{n\_neighbors} , \texttt{spread}, and \texttt{min\_dist}.  The \texttt{n\_neighbors} hyper-parameter ($K$) denotes the number of nearest neighbors to consider in constructing the high-dimensional fuzzy simplicial set for each datapoint.  \texttt{Spread} ($\gamma$) and \texttt{min\_dist} ($\delta$) function together to create the differentiable decay curve used to approximate the injective function for mapping between high- and low-dimensional fuzzy simplicial sets.  In essence, \texttt{min\_dist} determines the degree to which nearby points are clustered or expanded, relative to to the overall \texttt{spread}.

Two other parameters ultimately affect this balance between preserving local and global structure: the seed ($\tau$) for the (pseudo-)random number generator, and the order of the observations (i.e., trajectories) to be embedded ($\phi$).  As described in \textit{Methods}, we ensured the video and recall events were projected onto the \textit{same} low-dimensional manifold by fitting the embedding model to a stacked matrix of all video, average recall, and individuals' recall events.  After initializing the low-dimensional simplicial set (by default, using a spectral embedding of the high-dimensional simplicial set's fuzzy 1-skeleton), UMAP optimizes the embedding using stochastic gradient descent with cross-entropy as a cost function.  During optimization, indices of the data are sampled at random, and thus the local minimum achieved by the optimization is dependent on both the state of the RNG and the sequence in which observations are concatenated.

We performed a grid search over pre-specified values of these hyper-parameters, and optimized the manifold space for visualization based on two criteria: First, that the 2D embedding of the video trajectory should reflect its original 100-dimensional structure as faithfully as possible. Second, that the path traversed by the embedded video trajectory should intersect itself a minimal number of times.  The first criteria helps bolster the validity of visual intuitions about relationships between sections of video content, based on their locations in the embedding space.  The second criteria was motivated by the observed low off-diagonal values in the video trajectory's temporal correlation matrix (suggesting that the same topic-space coordinates should not be revisited; see Figure~2A in the main text) and the inability to accurately describe the consistency of participants' transitions through topic space at locations where those transitions frequently occurred in multiple, orthogonal directions (see \textit{Methods} for analysis details).  We formalized this optimization as the combination of hyper-parameter values that satisfied
\[
\argmax_{\left\{K, \gamma, \delta, \tau, \varphi~\in~E~\mathrm{|}~\Phi\left(K, \gamma, \delta, \tau, \varphi \right) \right\}} \left[\mathrm{corr}\left(\Upsilon\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right),~\Upsilon\left(\chi\right)\right)\right],~\mathrm{where}
\]
\[
\Phi\left(K, \gamma, \delta, \tau, \varphi \right) = \argmin_{K, \gamma, \delta, \tau, \varphi} \left[\Gamma\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right)\right],
\]
$\xi$ is the video's trajectory through the manifold space; $\chi$ is the original, 100-dimensional video trajectory; $\Upsilon$ is a condensed matrix of pairwise distances between event vectors (computed using correlation distance in the original topic space and Euclidean distance in the manifold space); $\mathrm{corr}\left(\Upsilon\left(\xi\right), \Upsilon\left(\chi\right)\right)$ is the correlation between the sets of pairwise distances, and $\Gamma$ is the number of instances in which lines drawn between consecutive video event embeddings intersected each other.   The sets of hyper-parameter values we searched over comprised: $K \in \{10x \in \mathbb{Z}~|~10 < x <  22\} \cup \{161\}$ (a range roughly centered on half the total number of events, 161, in order to balance representations of local and global structure); $\gamma \in \{1, 3, 5, 7, 9\}$; $\delta \in \{0.1, 0.3, 0.5, 0.7, 0.9\}$; $\tau \in \{x \in \mathbb{Z}~|~0 < x < 101\}$; and $\phi \in {S\choose3},~\mathrm{where}~S=\{\mathrm{video~events,~average~recall~events, individual~recall~events}\}$.  The optimal parameters (that yielded $\Phi=0$) were $K=170$, $\gamma=7$, $\delta=0.7$, $\tau=41$, with the order of sequence $\varphi$ as the average recall events, video events, and individuals' recall events, vertically concatenated, in order.
